---
title: "p8105_hw2_wz2590"
output: github_document
author: Weiheng Zhang
---



## Problem 1
```{r  message = FALSE}
library(tidyverse)
```


Read and clean the Mr. Trash Wheel sheet.

```{r message = FALSE, warning = FALSE}
trashWheels_df = readxl::read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                                    range = "A2:N534") %>%
  janitor::clean_names() %>% # make the naming uniform
  drop_na(dumpster) %>% # drop all the rows that is not dumpster specific
  mutate(sports_balls = round(sports_balls))

trashWheels_df
```

Read and clean precipitation data for 2018 and 2019.

```{r message = FALSE}
precipitation_18_raw = readxl::read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                                      sheet = "2018 Precipitation",
                                      range = "A2:B14")
#precipitation_18_raw

precipitation_18_tidy = precipitation_18_raw %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>%
  relocate(year)

precipitation_18_tidy
```

```{r message = FALSE}
precipitation_19_raw = readxl::read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                                      sheet = "2019 Precipitation",
                                      range = "A2:B14")
#precipitation_19_raw

precipitation_19_tidy = precipitation_19_raw %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019,) %>%
  relocate(year)

precipitation_19_tidy
```

Combine the 2018 and 2019 data, and convert month to a character variable
```{r message = FALSE}
precipitation_comb = 
  bind_rows(precipitation_18_tidy, precipitation_19_tidy) %>%
  mutate(month = month.name[month])

precipitation_comb
```


Calculate the median number of sports balls in a dumpster in 2019
```{r message = FALSE}
sportsBalls_2019 = trashWheels_df %>% 
  filter(year == 2019) %>% 
  select(sports_balls)

med_sportsBalls_2019 = median(sportsBalls_2019[[1]])

med_sportsBalls_2019
```






Mr. Trash Wheel is a water-wheel vessel that removes trash from the Inner Harbor in Baltimore, Maryland. There are `r nrow(trashWheels_df)` observations in the trashWheels_df data frame, each observation has `r ncol(trashWheels_df)` variables: Dumpster's label, Month, Year, Date, Weight_tons, volume_cubic_yards, plastic_bottles, polystyrene, cigarette_butts, glass_bottles, grocery_bags, chip_bags, sports_bags, homes_powered

The precipitation_18_tidy data frame contains the precipitation(in) for each month in year 2018. There are `r nrow(precipitation_18_tidy )` observations in this data frame, since a year has 12 months.

The precipitation_19_tidy data frame contains the precipitation(in) for each month in year 2019. There are `r nrow(precipitation_19_tidy )` observations in this data frame, since a year has 12 months.

The precipitation_comb data frame contains the precipitation(in) on each month from Jan 2018 to Dec 2019. There are `r nrow(precipitation_comb )` observations in this data frame, since two years have 24 months.

The key variables are `month` and `year`, because these variables exist in all data frames above and can be used to combine them. 

The total precipitation in 2018 is `r sum(precipitation_18_tidy$total)` inches. 
The median number of sports balls in a dumpster in 2019 is 9 balls.

  


## Problem 2


First, clean the data in pols-month.csv
```{r message = FALSE}
pols_month_df_raw = read_csv("./data/fivethirtyeight_datasets/pols-month.csv")

#pols_month_df_raw
 
pol_month_df = pols_month_df_raw %>%       
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), convert = T) %>% 
  mutate(month = month.name[month], 
         president = ifelse(prez_gop >= 1 & prez_dem == 0, "gop", "dem")) %>% 
  select(-c(day, prez_gop, prez_dem))
  
pol_month_df
```


Second, clean the data in snp.csv using a similar process to the above.
The format of the data before and after the year 2000 are different. Thus I need 
to read and tidy the old and new data separately. (It seems that Mac users don't have this problem?)

```{r message = FALSE}
snp_df_raw_new = read_csv("./data/fivethirtyeight_datasets/snp.csv", n_max = 175)

#snp_df_raw_old
 
snp_df_new = snp_df_raw_new %>%       
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month", "day"), convert = T) %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day)) %>% 
  arrange(year, month, day) %>% 
  mutate(month = month.name[as.numeric(month)], 
         year = year + 2000) 
  
#snp_df_new
```

```{r message = FALSE}
raw_colNames = read_csv("./data/fivethirtyeight_datasets/snp.csv", n_max = 0) %>% 
  names()

snp_df_raw_old = read_csv("./data/fivethirtyeight_datasets/snp.csv", skip = 176, col_names = raw_colNames)

#snp_df_raw_old
 
snp_df_old = snp_df_raw_old %>%       
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day)) %>% 
  arrange(year, month, day) %>% 
  mutate(month = month.name[as.numeric(month)], 
         year = ifelse(year == 00, year + 2000, year + 1900)) %>% 
  relocate(year, month, day, close)
  
  
#snp_df_old


snp_df = rbind(snp_df_old, snp_df_new) %>% 
  mutate(year = as.integer(year)) %>% 
  select(-day)
snp_df
```





Tidy the unemployment data so that it can be merged with the previous datasets. 

```{r message = FALSE}
unemployment_df_raw = read.csv("./data/fivethirtyeight_datasets/unemployment.csv") 

#unemployment_df_raw

unemployment_df = unemployment_df_raw %>%
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment_percentage") %>%
  mutate(month = match(month, month.abb), month = month.name[month]) %>% 
  janitor::clean_names()

unemployment_df
  
```


```{r message = FALSE}
merged_df = left_join(pol_month_df, snp_df, by = c("year", "month")) %>% 
  left_join(unemployment_df, by = c("year", "month"))

merged_df

year_range = merged_df %>% 
  select(year) %>% 
  range()

year_range
```



Describe the data:
The pol_month_df data frame describes the number of national politicians who are democratic or republican at any given time. It has `r nrow(pol_month_df)` observations, with each observations having `r ncol(pol_month_df)` variables.  
The snp_df data frame describes the Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. It has `r nrow(snp_df)` observations, with each observations having `r ncol(snp_df)` variables.  
The unemployment_df data frame contains the unemployment percentage for each month from January 1948 to December 2015. It has `r nrow(unemployment_df)` observations, with each observations having `r ncol(unemployment_df)` variables.  
The merged_df data frame is a combination of the previous three with key variables `year` and `month`. It has `r nrow(merged_df)` observations, with each observations having `r ncol(merged_df)` variables. The year ranges from 1947 to 2015.




## Problem 3


Load and tidy the data

```{r message = FALSE, warning = FALSE}
baby_names_df = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  #view %>% 
  mutate(ethnicity = recode(ethnicity, 
                            "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                            "BLACK NON HISP" = "BLACK NON HISPANIC",
                            "WHITE NON HISP" = "WHITE NON HISPANIC"),
         childs_first_name = str_to_sentence(childs_first_name)) %>% 
  distinct() %>% 
  view
```


Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time.

```{r message = FALSE}
Olivia_rankTable = baby_names_df %>% 
  filter(childs_first_name == "Olivia", gender == "FEMALE") %>%
  pivot_wider(names_from = "year_of_birth",
              values_from = "rank",
              id_cols = ethnicity
    )

Olivia_rankTable
  
```

Produce a similar table showing the most popular name among male children over time.

```{r message = FALSE}
boy_nameRank = baby_names_df %>% 
  filter(gender == "MALE", rank == 1) %>% 
  select(-count, -rank, -gender) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
    )
  
boy_nameRank
```


For male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r message = FALSE}
name_vs_rank_16 = baby_names_df %>% 
  filter(year_of_birth == 2016,
         gender == "MALE",
         ethnicity == "WHITE NON HISPANIC",
         ) %>%
  ggplot(aes(x = rank, y = count)) + geom_point() + 
  xlab("rank in popularity") + ylab("children name count") +
  labs(title = "Scatterplot of children name count vs. rank in popularity")

name_vs_rank_16
```

